---
title: "EDS 221 Day 3 Conditionals and Loops"
author: "Dalila Lara"
date: "2022-08-10"
output: html_document
---

### Setting global options to not print warning & messages

### The code chunk is running but the header states to not print code

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(tidyverse)
library(palmerpenguins)
```

# `if` statements

### Examples

```{r}
# making object
burrito <- 2.4 

# if burrito is greater than 2 print statement
if (burrito > 2) {
  print("I love burritos!")
}
```

### String example 
```{r}
# Creating vector 
my_ships <- c("Millenium Falcom", "X-wing", "Tie-Fighter", "Death Star")

# does the object contain the pattern "r"
str_detect(my_ships, pattern = "r")

phrase <- "I love burritos!"

if (str_detect(phrase, "love")) {
  print("I am a burrito fanl")
}
```

### Filtering data with filter,  Example of the %in% operator 
```{r}
# Starwars data
# view(starwars)
#Shortcut for pipe command + shift + m

#Creating subset of data with pipe and filter function
tatooine_characters <- starwars |> 
  filter(homeworld == "Tatooine")

# Use the %in% operator to check for multiple positive matches
tatooine_endor_naboo <- starwars |> 
  filter(homeworld %in% c("Tatooine", "Naboo", "Endor"))

# In words: Does the value in the home world variable for that row exsist IN that vector c("Tatooine", "Naboo", "Endor")
```

### `if-else` statement
```{r}
pika <- 50

if(pika > 60) {
  print("mega pika")
} else {
  print ("normal pika")
}
```

### `if-else-else if` statement
```{r}
marmot <- 2.8

if (marmot < 0.5) {
  print("this is a small marmot")
} else if (marmot >= 0.5 & marmot < 3) {
  print("this is a medium marmot")
} else {
  print("this is a large marmot")
}
```


# Lecture 6 Notes 

### `group_by() %>% summarize()
```{r}
#view(penguins)
# looks at penguins data, looks at species & island and calculates mean for each species/island group
# n is the sample size for the species as new row
penguins |>
  group_by(species, island) |>
  summarize(mean_bill_length = mean(bill_length_mm, na.rm = TRUE),
            sd_bill_length = sd(bill_length_mm, na.rm = TRUE),
            n = n())
unique(penguins$species)
```

### summarize(), across () and groupby() functions to find max/min etc of data 
### Making nested for loops with these built in functions
```{r}
# Unique function
unique(starwars$eye_color)

#1. Find the mean and sd of character height in 'starwars' grouped by homeworld
# Max and min of the homeworld for each charecter height
starwars |> 
  group_by(homeworld) |> 
  summarize(max_height = max(height, na.rm = TRUE),
            min_height = min(height, na.rm = TRUE))

#2. Find the maximum, minimum flipper length and sample size, for each species of penguins 
penguins |> 
  group_by(species, sex) |> 
  summarize(max_flipper = max(flipper_length_mm, na.rm = TRUE),
            min_flipper = min(flipper_length_mm, na.rm = TRUE))

# 3. Takes the mean value of speicies for any column that is numeric
penguins |> 
  group_by(species) |> 
  summarize(across(where(is.numeric), mean, na.rm = TRUE))
```

# Computational Notes 
### For loops
```{r}
# Create vector
tree_height <- c(1, 2, 6, 10, 14, 20)

# Loop, adding each element to the next
for (i in seq_along(tree_height)) {
  val = tree_height[i] + tree_height[i + 1]
  print(val)
}
```

### `if else else if` statement 
```{r}
# Vector that converts human years into animal years although the vector differs depending on animal

species <- c("dog", "elephant", "goat", "dog", "dog", "elephant")
age_human <- c(3, 8, 4, 6, 12, 18)

# empty vector to hold converted animal ages in human ages
animal_ages <- vector(mode = "numeric", length = length(species))
for (i in seq_along(species)) {
  if (species[i] == "dog") {
    animal_age <- age_human[i] * 7
  } else if(species[i] == "elephant") {
    animal_age <- age_human[i] * 0.88
  } else if(species[i] == "goat") {
    animal_age <-age_human[i] * 4.7
  } 
  animal_ages[i] <- animal_age
}
```

### Example 2 w/mtcars
```{r}
#view(mtcars)

mean_mtcars <- vector(mode = "numeric", length = ncol(mtcars))

for (i in seq_along(mtcars)) {
  mean_val <- mean(mtcars[[i]])
}

mean_mtcars
```

### Functional programming examples 
Adding  syntactic sugar (apply, cross, map_df)
```{r}
apply(X = mtcars, MARGIN = 2, FUN = mean)
```

```{r}
map(.x = mtcars, .f = mean)
```





